An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-23 EconGrowth.Rnw
Edited the to-do list slightly.
Now including parameter graphs for CES models. Need to adjust vertical axes and tic marks.

\section*{To-do List}
\begin{itemize}
\item Reese: Add text from Word verson of paper to the LaTeX version.
\item Heun: Add CES production function fits
      \begin{itemize}
      \item \st{GDP comparison graph}
      \item Convert $\rho$ variables in the \citet{Henningsen:2011td} CES production function to $\sigma$ variables.
      \item parameter graphs for each country
      \item parameter tables for each country in the appendix
      \item AIC rows in the AIC table
      \end{itemize}
\item Heun: Deal with missing covariance numbers when columns have unequal length. Use the \texttt{subset} command.
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the Cobb-Douglas with energy rows. Asked Pruim about this via email on 4 Feb 2013 but have not heard a response.
      \begin{itemize}
      \item ZA: lower bound on $\lambda$ and upper bound on $\alpha$
      \item ZM: lower bound on $\lambda$
      \end{itemize}
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the LINEX model
      \begin{itemize}
      \item Zambia, upper bound on $c_t$
      \end{itemize}
\item Heun: Fix warnings of the form ``Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562'' in the code that generates the Cobb-Douglas with energy fits.
\item Heun: x-axis labels only on bottom for 9-panel lattice graphs
\item Heun: adjust y axis limits on $a_0$ and $c_t$ parameter graphs for the LINEX model AFTER we finish debugging the LINEX models.
\item de Wit: provide funding acknowledgement, if necessary.
\item de Wit: review, critique, and contextualize.
\item \st{Heun: Ask Martin if he can supply an ``Economics'' affiliation. Will probably make a stronger paper.}
\item \st{Heun: Create a table of AIC values for each fit.}
\item \st{Heun: Add $u$ predictions for Cobb-Douglas}
\item \st{Heun: Move tables to an appendix at the end.}
\item \st{Heun: Add acknowledgements (Dad, Pruim)}
\item \st{Heun: Add $u$ parameter table and graph}
\item \st{Heun: Add covariance metrics.}
\item \st{Heun: Add single factor fits}
      \begin{itemize}
      \item \st{Correct sizes for parameter graphs for each country}
      \item \st{Parameter graphs ($\lambda$ and $m$) for each country}
      \item \st{Parameter tables for each country in the appendix}
      \item \st{GDP comparison graph}
      \item \st{AIC rows in the AIC table}
      \end{itemize}
\item \st{Heun: Add setup variables for figure height and width where not already included.}
\item \st{Heun: Add LINEX production function fits}
      \begin{itemize}
      \item \st{Parameter graphs for each country}
      \item \st{Parameter tables for each country in the appendix}
      \item \st{AIC rows in the AIC table}
      \item \st{GDP comparison graph}
      \end{itemize}
\end{itemize}

@@

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countryAbbrevs <- c(US="US", UK="UK", JP="JP", CN="CN", ZA="ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
countryAbbrevsU <- c(US="US", UK="UK", JP="JP") #Only these countries have useful work data
countryNamesAlph <- c(CN="China", IR="Iran", JP="Japan", SA="Saudi Arabia", TZ="Tanzania", UK="United Kingdom", US="USA", ZA="South Africa", ZM="Zambia") #In alphabetical order.
energyTypes <- c(Q="Q", X="X", U="U") # List of energy types
factors <- c(K="K", L="L", Q="Q", X="X", U="U") # List of factors of production
########### Several global parameters for graphs. Set here and use below to ensure consistent appearance of graphs.
# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. 
# "1" means first alphabetically.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9) # Sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
maxWidth <- 6.5 #Inches
# Full page lattice plot sizes
ninePanelLayoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
ninePanelGraphWidth <- maxWidth     
ninePanelGraphHeight <- 7.5  #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
keyXLoc <- 0.01 #x position of the key
keyYLoc <- 0.95 #y position of the key
# Parameter graphs are tricky to ensure same size. I've found that with height set to 5 in for the 
# cd graphs and 2.93 in for the sf graphs, I get a nearly-exact match with graph size. From these settings I 
# can set up 2 equations:
#    5 in = 2*hgraph + haxis
#    2.93 in = hgraph + haxis
# to estimate that h_axis = 0.86 in and hgraph = 2.05 in.
h_param_graph <- 2.07 # Inches for a single panel
h_param_axis <- 0.86 # Inches (also includes whatever margin is applied)
sfParameterGraphHeight <- h_param_graph + h_param_axis
cdParameterGraphHeight <- 2*h_param_graph + h_param_axis
linexParameterGraphHeight <- h_param_graph + h_param_axis
sfParameterGraphWidth <- maxWidth
cdParameterGraphWidth <- maxWidth
cesParameterGraphWidth <- maxWidth
linexParameterGraphWidth <- maxWidth
# Other graph parameters that apply to all graphs
scaleTextSize <- 1.0  #Multiple of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
@

@@

#########################
# Takes the Cobb-Douglas fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
cobbDouglasPredictions <- function(countryAbbrev, energyType){
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!missing(energyType)){
    # Consider this replacement only if energyType has been specified.
    if (!(haveDataCD(countryAbbrev, energyType))){
      #If we don't have data for this combination of countryAbbrev and energyType, 
      # return a column of NAs if the above conditions have been met.
      nRows <- 21 # All of these countries need 21 rows.
      df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
      colnames(df) <- "pred"
      return(df)
    }
  }
  model <- cobbDouglasModel(countryAbbrev, energyType)
  pred <- predict(model) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

@@

@@

@@

@@

@@

<<CES_Params_Table, results='asis', eval=TRUE>>=
printCESParamsTable("Q"); printCESParamsTable("X"); printCESParamsTable("U")
@

>>>>

\section*{To-do List}
\begin{itemize}
\item Reese: Add text from Word verson of paper to the LaTeX version.
\item Heun: Add CES production function fits
      \begin{itemize}
      \item \st{GDP comparison graph}
      \item Convert $\rho$ variables in the \citet{Henningsen:2011td} CES production function to $\sigma$ variables.
      \item Create parameter graphs for each country
      \item \st{Parameter tables for each energy type in the appendix}
      \item Fix the problems with $\sigma_1$ and $\sigma$ values in the CES parameter tables in the appendix
      \item AIC rows in the AIC table
      \end{itemize}
\item Heun: Deal with missing covariance numbers when columns have unequal length. Use the \texttt{subset} command.
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the Cobb-Douglas with energy rows. Asked Pruim about this via email on 4 Feb 2013 but have not heard a response.
      \begin{itemize}
      \item ZA: lower bound on $\lambda$ and upper bound on $\alpha$
      \item ZM: lower bound on $\lambda$
      \end{itemize}
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the LINEX model
      \begin{itemize}
      \item Zambia, upper bound on $c_t$
      \end{itemize}
\item Heun: Fix warnings of the form ``Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562'' in the code that generates the Cobb-Douglas with energy fits.
\item Heun: x-axis labels only on bottom for 9-panel lattice graphs
\item Heun: adjust y axis limits on $a_0$ and $c_t$ parameter graphs for the LINEX model AFTER we finish debugging the LINEX models.
\item de Wit: provide funding acknowledgement, if necessary.
\item de Wit: review, critique, and contextualize.
\item \st{Heun: Ask Martin if he can supply an ``Economics'' affiliation. Will probably make a stronger paper.}
\item \st{Heun: Create a table of AIC values for each fit.}
\item \st{Heun: Add $u$ predictions for Cobb-Douglas}
\item \st{Heun: Move tables to an appendix at the end.}
\item \st{Heun: Add acknowledgements (Dad, Pruim)}
\item \st{Heun: Add $u$ parameter table and graph}
\item \st{Heun: Add covariance metrics.}
\item \st{Heun: Add single factor fits}
      \begin{itemize}
      \item \st{Correct sizes for parameter graphs for each country}
      \item \st{Parameter graphs ($\lambda$ and $m$) for each country}
      \item \st{Parameter tables for each country in the appendix}
      \item \st{GDP comparison graph}
      \item \st{AIC rows in the AIC table}
      \end{itemize}
\item \st{Heun: Add setup variables for figure height and width where not already included.}
\item \st{Heun: Add LINEX production function fits}
      \begin{itemize}
      \item \st{Parameter graphs for each country}
      \item \st{Parameter tables for each country in the appendix}
      \item \st{AIC rows in the AIC table}
      \item \st{GDP comparison graph}
      \end{itemize}
\end{itemize}

@@

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countryAbbrevs <- c(US="US", UK="UK", JP="JP", CN="CN", ZA="ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
countryAbbrevsU <- c(US="US", UK="UK", JP="JP") #Only these countries have useful work data
countryNamesAlph <- c(CN="China", IR="Iran", JP="Japan", SA="Saudi Arabia", TZ="Tanzania", UK="United Kingdom", US="USA", ZA="South Africa", ZM="Zambia") #In alphabetical order.
energyTypes <- c(Q="Q", X="X", U="U") # List of energy types
factors <- c(K="K", L="L", Q="Q", X="X", U="U") # List of factors of production
########### Several global parameters for graphs. Set here and use below to ensure consistent appearance of graphs.
# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. 
# "1" means first alphabetically.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9) # Sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
maxWidth <- 6.5 #Inches
# Full page lattice plot sizes
ninePanelLayoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
ninePanelGraphWidth <- maxWidth     
ninePanelGraphHeight <- 7.5  #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
keyXLoc <- 0.01 #x position of the key
keyYLoc <- 0.95 #y position of the key
# Parameter graphs are tricky to ensure same size. I've found that with height set to 5 in for the 
# cd graphs and 2.93 in for the sf graphs, I get a nearly-exact match with graph size. From these settings I 
# can set up 2 equations:
#    5 in = 2*hgraph + haxis
#    2.93 in = hgraph + haxis
# to estimate that h_axis = 0.86 in and hgraph = 2.05 in.
h_param_graph <- 2.07 # Inches for a single panel
h_param_axis <- 0.86 # Inches (also includes whatever margin is applied)
sfParameterGraphHeight <- h_param_graph + h_param_axis
cdParameterGraphHeight <- 2*h_param_graph + h_param_axis
cesParameterGraphHeight <- 3*h_param_graph + h_param_axis
linexParameterGraphHeight <- h_param_graph + h_param_axis
sfParameterGraphWidth <- maxWidth
cdParameterGraphWidth <- maxWidth
cesParameterGraphWidth <- maxWidth
linexParameterGraphWidth <- maxWidth
# Other graph parameters that apply to all graphs
scaleTextSize <- 1.0  #Multiple of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
@

@@

# print("cobbDouglasModel(US, Q) works fine:")
# us <- cobbDouglasModel("US", "Q")
# print(summary(us))
# print(confint(us))
# print("cobbDouglasModel(ZA, Q) has NA values:")
# za <- cobbDouglasModel("ZA", "Q")
# print(summary(za))
# print(confint(za))
# save(us, za, file="HeunModels.Rdata")

#########################
# Takes the Cobb-Douglas fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
cobbDouglasPredictions <- function(countryAbbrev, energyType){
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!missing(energyType)){
    # Consider this replacement only if energyType has been specified.
    if (!(haveDataCD(countryAbbrev, energyType))){
      # If we don't have data for this combination of countryAbbrev and energyType, 
      # return a column of NAs when the above conditions have been met.
      nRows <- 21 # All of these countries need 21 rows.
      df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
      colnames(df) <- "pred"
      return(df)
    }
  }
  model <- cobbDouglasModel(countryAbbrev, energyType)
  pred <- predict(model) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

@@

#############################
# Creates a graph with confidence intervals for the CES model for the given energyType
##
createCESParamsGraph <- function(energyType){
  # Create a data table with the following columns:
  # country abbrev, parameter (gamma, lambda, delta_1, delta, sigma_1, sigma), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, cesCountryRowsForParamsGraph, energyType=energyType))
  graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                   data = dataTable, 
                   centers = value, #identifies where the dots should be placed
                   as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                   draw.bands = FALSE, #provides nicer error bars
                   horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                   layout = c(2,3), #2 column, 3 row
                   # Orders the panels as gamma, lambda, delta_1, delta, sigma_1, sigma
                   index.cond = list(c(1,2,3,4,5,6)), 
                   #set labels and bg color in strip
                   strip = strip.custom(factor.levels=c("$\\gamma$", "$\\lambda$", "$\\delta_1$",
                                                        "$\\delta$", "$\\sigma_1$", "$\\sigma$"), bg="white"),
                   col = "black", #Sets line color to black
                   lwd = 1, #Sets line width to 1.0
                   ylim = list(c(-0.05, 1.0), c(-0.5, 5), c(0.0, 5.0),
                               c(0.0, 1.0), c(0.0, 1.0), c(0.0, 1.0)), #y axis limits
                   scales = list(cex=scaleTextSize, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 x=list(cex=0.75), #reduces text size so that country abbrevs are legible
                                 y=list(relation="free", #allow each axis to be different
                                        at=list(c(0.0, 0.5, 1.0), c(0, 1, 2, 3, 4, 5), c(0.0, 0.5, 1.0),
                                                c(0.0, 0.5, 1.0), c(0.0, 0.5, 1.0), c(0.0, 0.5, 1.0))   #y tick marks
                                       )
                   )
  )
  return(graph)
}

@@

Figure \ref{fig:CESq_Params_Graph} shows values and 95\% confidence intervals for the parameters for the CES model (with $q$).

<<CESq_Params_Graph, fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.pos="H", fig.cap="CES (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createCESParamsGraph("Q")
@

@@

Figure \ref{fig:CESx_Params_Graph} shows values and 95\% confidence intervals for the parameters for the CES model (with $x$).

<<CESx_Params_Graph, fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.pos="H", fig.cap="CES (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createCESParamsGraph("X")
@

@@

Figure \ref{fig:CESu_Params_Graph} shows values and 95\% confidence intervals for the parameters for the CES model (with $u$).

<<CESu_Params_Graph, fig.width=cesParameterGraphWidth, fig.height=cesParameterGraphHeight, fig.pos="H", fig.cap="CES (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createCESParamsGraph("U")
@

@@

<<CES_Params_Table, results='asis'>>=
printCESParamsTable("Q"); printCESParamsTable("X"); printCESParamsTable("U")
@
